---
title: "CS 422 Section 01"
output: html_notebook
author: Zhongzhu Ding
---

## Homework 3(p1)

Perform K-means clustering on file19.txt on the above web page. This file contains a multivariate mammals
dataset; there are 9 columns and 66 rows.
(a) Data cleanup
(i) Think of what attributes, if any, you may want to omit from the dataset when you do the clustering. Indicate all of the attributes you removed before doing the clustering.
(ii) Does the data need to be standardized?
(iii) You will have to clean the data to remove multiple spaces and make the comma character the delimiter.
Please make sure you include your cleaned dataset in the archive file you upload

### Part 2.1-A-i

```{r}
library(cluster)
setwd("/users/enrique/desktop/IITstudy/CS422 Data mining/hw/hw3")
#skip straight to the table
mammals <- read.table("file19.txt",skip =20,header = T) 
#omit the attributes from the datasets
mammals.cp <- mammals[,-c(5,7,9)]

```
(i)I omit the attributes c,p and m from the dataset because they got less difference in the tooth distribution than the first group (I,i)

### Part 2.1-A-ii
(ii)No, it doesn't need to be standardized because the attributes's value are in the same small ranges.

### Part 2.1-A-iii
```{r}
#remove multiple spaces, make the swlimiter and output.
mammals.cp$Name <- trimws(mammals.cp$Name)
write.table(mammals.cp,"mammals.csv",sep = ",") 
```

(iii)the cleaned dataset is included in the archive file.


b) Clustering
(i) Determine how many clusters are needed by running the WSS or Silhouette graph. Plot the graph using
fviz_nbclust().
(ii) Once you have determined the number of clusters, run k-means clustering on the dataset to create that many clusters. Plot the clusters using fviz_cluster().
(iii) How many observations are in each cluster?
(iv) What is the total SSE of the clusters?
(v) What is the SSE of each cluster?
(vi) Perform an analysis of each cluster to determine how the mammals are grouped in each cluster, and whether that makes sense?
assuming k is the variable that holds the output of the kmeans() function call.

### Part 2.1-B-i
```{r}
library(factoextra)
fviz_nbclust(mammals.cp[2:6],kmeans,method="wss")
fviz_nbclust(mammals.cp[2:6],kmeans,method="silhouette")

```

(i)Considering both of the WSS and Silhouette graphs,we can see that 10 clusters is the best choice.

### Part 2.1-B-ii
```{r}
k <- kmeans(mammals.cp[2:6],centers = 10,nstart = 25)

fviz_cluster(k,data = mammals.cp[2:6])
```

### Part 2.1-B-iii
```{r}
k
```

(iii)the number of observations from cluster1 to cluster10 are 7, 8, 11, 1, 11, 9, 2, 1, 10, 6.

### Part 2.1-B-(iv)
```{r}
bd = mammals.cp[,-1]
wss = kmeans(bd, centers=1)$tot.withinss

for (i in 2:10)
  wss[i] = kmeans(bd, centers=i)$tot.withinss

library(ggvis)
sse = data.frame(c(1:10), c(wss))
names(sse)[1] = 'Clusters'
names(sse)[2] = 'SSE'
sse %>%
  ggvis(~Clusters, ~SSE) %>%
  layer_points(fill := 'blue') %>% 
  layer_lines() %>%
  set_options(height = 300, width = 400)
```
### Part 2.1-B-(iv)
```{r}
sum(sse)
```

(iv)the SSEtotal=SSE1+SSE2+...+SSE10=1242.768

### Part 2.1-B-(v)
```{r}
sse
```

(v)the SSE in each cluster are listed in the diagram above.

### Part 2.1-B-(vi)
```{r}
which(k$cluster == 1)
which(k$cluster == 2)
which(k$cluster == 3)
which(k$cluster == 4)
which(k$cluster == 5)
which(k$cluster == 6)
which(k$cluster == 7)
which(k$cluster == 8)
which(k$cluster == 9)
which(k$cluster == 10)
```
(vi)Taking cluster1 as an example,the indices of the animals are [2 3 4 5 32 37 57],which refers to Hairy tail mole,Common mole,Star nose mole,Brown bat,Coyote,Raccoon and Peccary.All of number distribution in attributes are pretty similar.Just as the example of cluster1,by comparing the number in 5 attributes,the mammals' grouping makes sense.
